---
title: "Machine learning project for data science"
output: html_document
---
##Introduction
This analysis aims to build a predictive model for the Weight Lifting Exercises Dataset.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load R packages and download data file
First, download and read the files. Missing values or error values are set to NA.

```{r}
library(data.table)
library(ggplot2)
library(lattice)
library(caret)
library(randomForest)
library(rpart)
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","training.csv")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv","testing.csv")
trainingfile<-fread("/Users/daxinggao/training.csv", na.strings=c("NA","#DIV/0!",""))
testingfile<-fread("/Users/daxinggao/testing.csv", na.strings=c("NA","#DIV/0!",""))
```

## Preprocess data
Create training and testing sets from training files. Only consider the columns that have fewer than 60% NA. And the first seven columns are also discarded.

```{r}
set.seed(1234)
inTrain<-createDataPartition(y=trainingfile$classe,p=0.75,list=F)#create training and testing sets.
training<-trainingfile[inTrain]
testing<-trainingfile[-inTrain]
NAcol<-sapply(training,function(x){sum(is.na(x))/nrow(training)})#calculate number of NA for each column
NAcolN<-match(names(NAcol[NAcol>=0.6]),colnames(training))#columns have more than 60% NA
preprocess<-preProcess(training[,-c(1:7,NAcolN,ncol(training)),with=F],method=c("center","scale","knnImpute"))#preprocess data, discard first seven columns/columns have more than 60% NA/last column, impute missing data with knnImpute
trainx<-predict(preprocess,training[,-c(1:7,NAcolN,ncol(training)),with=F])
trainnew<-cbind(trainx,training[,ncol(training),with=F])
```

## Use random forest model to fit and calculate prediction accuracy
```{r}

model<-randomForest(factor(classe) ~ .,trainnew)
confusionMatrix(predict(model,trainx),trainnew$classe)
```

## Preprocess and predict the testing data and calculate prediction accuracy


```{r}
testx<-predict(preprocess,testing[,-c(1:7,NAcolN,ncol(testing)),with=F])
testnew<-cbind(testx,testing[,ncol(testing),with=F])
confusionMatrix(predict(model,testx),testnew$classe)
```
##Predict data from testing file
```{r}
testfilex<-predict(preprocess,testingfile[,-c(1:7,NAcolN,ncol(testingfile)),with=F])
predict(model,testfilex)
```

##Conlusion
The random forest model successfully predicts the Weight Lifting Exercises Dataset.